\chapter{Language implementation}

\section{Lexer Implementation}

The lexer is responsible for transforming raw source text into a flat sequence of tokens, which the parser then consumes. It operates line by line, handling both the structural indentation logic and the inline tokenization of each line's content.

\subsection{Token Recognition}

Token recognition is driven by a list of ordered token specifications, each pairing a \texttt{TokenType} with a regular expression pattern. The order is significant --- more specific or longer patterns are listed before shorter ones to prevent incorrect partial matches. For example, the two-character operators \texttt{!=}, \texttt{==}, \texttt{<=}, and \texttt{>=} are specified before their single-character counterparts \texttt{<} and \texttt{>}.

All individual patterns are combined at class initialization into a single master regular expression using alternation, and compiled once into a \texttt{Pattern} object. During tokenization, a \texttt{Matcher} is advanced through the content of each line using \texttt{lookingAt()}, which attempts to match at the current position without requiring a full-string match. The capturing group that produced the match is then used to identify the corresponding \texttt{TokenSpec}.

\subsection{Indentation Handling}

Since the DSL uses indentation-based block delimitation, the lexer is responsible for emitting \texttt{INDENT} and \texttt{DEDENT} tokens explicitly. 
An integer stack tracks the indentation levels encountered so far, initialized with a base level of zero. At the start of each line, the number of leading spaces and tabs is counted. 
If the measured indentation exceeds the top of the stack, an \texttt{INDENT} token is emitted and the new level is pushed. If it is less, \texttt{DEDENT} tokens are emitted and levels are popped until the stack matches the current indentation. 
Blank and whitespace-only lines are skipped entirely and carry no structural meaning.

\subsection{Keyword Resolution and Special Cases}

Identifiers are first matched by a general alphanumeric pattern, then passed through a keyword lookup to determine whether the matched text corresponds to a reserved word such as \texttt{for}, \texttt{while}, \texttt{if}, or \texttt{true}. If a keyword match is found, the corresponding keyword token type is emitted instead of a generic \texttt{IDENT} token.

String literals have their surrounding quotation marks stripped during tokenization, so the parser receives only the inner content. Comments, introduced by a double dash (\texttt{--}), cause the lexer to discard the remainder of the line immediately upon detection. Any character that does not match any known pattern results in an \texttt{ILLEGAL} token carrying the unrecognized character, allowing the parser to produce a meaningful error message rather than failing silently.

Each line concludes with an explicit \texttt{NEWLINE} token. At the end of the source, any remaining open indentation levels are closed with corresponding \texttt{DEDENT} tokens, followed by a final \texttt{EOF} token.

\section{Parser Implementation}

The parser consumes the flat token sequence produced by the lexer and constructs an Abstract Syntax Tree (AST), which represents the hierarchical structure of the program. It is implemented as a hand-written recursive descent parser, where each grammar rule corresponds directly to a parsing method.

\subsection{Token Navigation}

The parser maintains a cursor over the token list and exposes three core navigation primitives. The \texttt{advance()} method moves the cursor forward and returns the new current token. The \texttt{check()} method tests whether the current token matches a given type without consuming it. The \texttt{expect()} method asserts that the current token has a specific type, consumes it, and throws a \texttt{ParseException} with a descriptive message including the line and column number if the assertion fails. A \texttt{peek()} method allows looking ahead by an arbitrary offset, which is used for disambiguation without side effects.

\subsection{Statement Parsing and Disambiguation}

The top-level entry point, \texttt{parseProgram()}, skips any leading newlines and delegates to \texttt{parseStatementList()}, which iterates until an \texttt{EOF} or, when parsing inside a block, a \texttt{DEDENT} token is encountered.

Individual statements are dispatched in \texttt{parseStatement()} based on the type of the current token. Control flow keywords (\texttt{for}, \texttt{while}, \texttt{if}) are routed directly to their respective parsing methods. When the current token is an identifier, a one-token lookahead is used to disambiguate between an assignment, where the next token is \texttt{=}, and a function call statement, where the next token is \texttt{(}. Any identifier-headed expression that resolves to something other than a function call is rejected with a parse error, enforcing the grammar's restriction that bare expressions are not valid statements.

\subsection{Control Flow Constructs}

The two forms of \texttt{for} loop are disambiguated after consuming the \texttt{for} keyword by checking whether the next token is an identifier followed by the \texttt{from} keyword. If so, a range loop is parsed; otherwise, a count loop is assumed. Both forms conclude by parsing an indented block via \texttt{parseBlock()}, which expects an \texttt{INDENT} token, recursively parses a statement list, and then expects a \texttt{DEDENT} token. The \texttt{if} statement optionally parses an \texttt{else} branch by checking for the \texttt{else} keyword after the then-block, with intervening newlines skipped.

\subsection{Expression and Condition Parsing}

Expressions are parsed using a standard two-level recursive descent that naturally encodes operator precedence. The \texttt{parseExpression()} method handles additive operators by parsing a \texttt{parseTerm()} and then consuming any sequence of \texttt{+} or \texttt{-} operators followed by additional terms, constructing left-associative \texttt{BinaryOp} nodes. The \texttt{parseTerm()} method handles multiplicative operators in the same manner, and \texttt{parseFactor()} handles the base cases: numeric, string, and boolean literals; parenthesised expressions; unary negation; and identifier-headed constructs.

When an identifier is encountered in \texttt{parseFactor()}, a further one-token lookahead determines whether it begins a function call (\texttt{IDENT} followed by \texttt{(}), a dot-notation special object access (\texttt{IDENT} followed by \texttt{.}), or a plain variable reference.

Conditions are parsed separately from expressions. The \texttt{parseCondition()} method first checks for a \texttt{not} keyword, in which case it recursively parses a negated condition. Otherwise, it parses a left-hand expression and checks whether the current token is a comparison operator. If so, a \texttt{ComparisonCondition} node is produced; if not, the expression itself is wrapped in a \texttt{BooleanCondition}, supporting the use of bare identifiers and function calls as boolean predicates.

\subsection{Error Handling}

All syntax errors are reported by throwing a \texttt{ParseException}, an unchecked exception that carries a human-readable message including the offending token type, its literal value where applicable, and its line and column position. This ensures that error messages are actionable for the learner rather than presenting raw internal state.