\chapter{Project Plan}

This appendix presents the project schedule, deliverables, and team responsibilities
across the full 15-week development cycle. The project is structured into four
sequential phases, each corresponding to a distinct stage of the software
engineering process: problem identification and research, language planning and
grammar design, core development and milestone evaluation, and final delivery.

\section*{Development Phases}

\begin{description}
  \item[Phase 1 --- Problem \& Research (Weeks 1--3)]
    Covers the initial definition of the problem domain, the construction of the
    PBL Problem Map, and a systematic literature review comprising a minimum of
    twenty references.

  \item[Phase 2 --- Planning \& Grammar (Weeks 4--6)]
    Encompasses the preparation of the project plan, the formal specification of
    the language grammar in Backus--Naur Form, and the implementation of the
    lexer, parser, and Abstract Syntax Tree.

  \item[Phase 3 --- Development \& Milestones (Weeks 7--13)]
    Covers the writing of the scientific paper abstract and draft, the Mid-term~I
    and Mid-term~II presentations, and the completion of the frontend
    implementation.

  \item[Phase 4 --- Final Delivery (Weeks 14--15)]
    Comprises the submission of the final report draft and the delivery of the
    final project presentation before the full team and evaluators.
\end{description}

\section*{Deliverables Timeline and Responsibilities}

The table below enumerates all thirteen project deliverables, their associated
week, and the team member responsible for each.

\bigskip

\ifdefined\colA\else\newlength{\colA}\fi
\ifdefined\colB\else\newlength{\colB}\fi
\ifdefined\colC\else\newlength{\colC}\fi
\setlength{\colA}{\dimexpr 0.10\textwidth - \tabcolsep\relax}
\setlength{\colB}{\dimexpr 0.63\textwidth - \tabcolsep\relax}
\setlength{\colC}{\dimexpr 0.27\textwidth - 2\tabcolsep\relax}

\noindent\begin{tabular}{@{}
  >{\raggedright\arraybackslash}p{\colA}
  >{\raggedright\arraybackslash}p{\colB}
  >{\raggedright\arraybackslash}p{\colC}@{}}
\toprule
\textbf{Week} & \textbf{Deliverable} & \textbf{Responsible} \\
\midrule
\multicolumn{3}{@{}l}{\textit{Phase 1 --- Problem \& Research}} \\[3pt]
W1  & Domain selection and problem identification      & Member 2        \\[2pt]
W2  & PBL Problem Map and domain analysis              & Member 5        \\[2pt]
W3  & Literature analysis (minimum 20 references)      & Member 5        \\[6pt]
\multicolumn{3}{@{}l}{\textit{Phase 2 --- Planning \& Grammar}} \\[3pt]
W4  & Project plan                                     & Members 1, 5    \\[2pt]
W5  & Formal grammar specification (BNF)               & Member 2        \\[2pt]
W6  & Lexer, parser, and Abstract Syntax Tree          & Member 1        \\[6pt]
\multicolumn{3}{@{}l}{\textit{Phase 3 --- Development \& Milestones}} \\[3pt]
W7  & Scientific paper abstract                        & Member 5        \\[2pt]
W8--9  & Mid-term I presentation                       & Member 5        \\[2pt]
W10--11 & Frontend implementation complete             & Member 1        \\[2pt]
W12 & Scientific paper draft                           & Member 5        \\[2pt]
W13 & Technical demonstration (Mid-term II)            & Member 4        \\[6pt]
\multicolumn{3}{@{}l}{\textit{Phase 4 --- Final Delivery}} \\[3pt]
W14 & Final report draft                               & Member 5        \\[2pt]
W15 & Final presentation                               & All Members     \\
\bottomrule
\end{tabular}

\bigskip
\noindent\textit{Total: 15 weeks, 13 deliverables, 5 team members.}

\clearpage
